---
title: "Unsupervised Learning"
author: "M.Erim Çelen"
date : "02 2020"
output: html_document
---
## Association Rules


In order to find Associations among products we will be using the groceries data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, eval=FALSE, include=FALSE}
install.packages("arules")
install.packages("arulesViz")
```

```{r include=FALSE}
library(arules)
library(arulesViz)
```

Let's load the groceries data.
```{r}
data(Groceries)
```
Groceries data has 9835 transactions and 169 items. There are 33 columns in the data set.

Let's explore the data.

```{r}
summary(Groceries)
```
We can see that the most items in the dataset are:
Whole milk
Other vegatables
Rolls/buns
Soda
Yogurt

```{r}
inspect(Groceries[1:5])
```

```{r}
LIST(head(Groceries))
```

```{r}
itemFrequencyPlot(Groceries,topN=20,type="absolute")
```

After looking at the data we can apply Apriori algorithm to data set.. 

Let's say we want to find out which items are sold three times a day, our support level will be 90/9835=0.009.


```{r}
rules <- apriori(Groceries, parameter=list(supp=0.009, conf=0.35)) 
```
We can see that there are 120 rules have been generated.

```{r}
summary(rules)
```

Summary functions show us 2 items are in 52 rules and 3 items are in 68 rules.

```{r}
rules.sup <- sort(rules, by="support", decreasing=TRUE) 
inspect(rules.sup[1:5])
```

We can also see these rules by lift, count and confidence.

The highest support level belongs to other vegetables let's see what forces make people buying other vegetables.
```{r}
rules.othervegetables <- apriori(data=Groceries, parameter=list(supp = 0.009,conf = 0.30), 
appearance=list(default = "lhs", rhs = "other vegetables"), control = list(verbose=F)) 

rules.othervegetables.byconf <- sort(rules.othervegetables, by ="confidence", decreasing = TRUE)
inspect(head(rules.othervegetables.byconf))
```
and oppositely if we want to see if we have other vegetables in a basket what more do we buy?
```{r}
rules.othervegetables <- apriori(data=Groceries, parameter=list(supp=0.009,conf = 0.20), 
appearance=list(default="rhs",lhs="other vegetables"), control=list(verbose=F)) 

rules.othervegetables.byconf<-sort(rules.othervegetables, by="confidence", decreasing=TRUE)
inspect(head(rules.othervegetables.byconf))
```


```{r}
plot(rules, method="grouped")  
```

We can also apply Apriori algorithm to closed frequent itemsets.

```{r}
groceries.closed <- apriori(Groceries, parameter = list(target = "closed frequent itemsets", support=0.10))

```


```{r}
inspect(groceries.closed)
```

```{r}
class(groceries.closed)
```

```{r}
is.closed(groceries.closed)  
```

We can also check the significance of the algorithm.
 
```{r}
is.significant(rules.othervegetables, Groceries)
```


```{r}
is.maximal(rules.othervegetables)
```


```{r}
is.redundant(rules.othervegetables)
```

```{r}
inspect(rules.othervegetables[is.redundant(rules.othervegetables)==FALSE])
```


We can also find Superset and subset in transactions.

```{r}
is.subset(rules.othervegetables, sparse = FALSE)
```


```{r}
is.superset(rules.othervegetables, sparse = FALSE)
```

Jaccard Index

```{r}
jaccard <- Groceries[,itemFrequency(Groceries)>0.1]

dissimilarity(jaccard, which="items") 
```

We can interpret these results as water and soda are very dissimilar in %89 they don't overlap.
